Pythran: Turning Python Modules into Optimized C++ Meta-Programs

### 2013-10-02 doc/papers/cc2014/paper.tex
Take Into account Pierrick's remark

% we don't say we focus on scientifical python?
% Ex : ...and most of \ac{OO} features to focus on scientific python modules
Pythran\cite{pythran2013} is an \ac{AOT} compiler that focuses on the
optimization of the functional and procedural paradigms offered by Python,
leaving aside dynamicity and most of \ac{OO} features.  Indeed, the Python
language and its standard library offers comprehensive services to the
programmers, but the language proves to be useful even with a limited subset of
features:
\begin{inparaenum}[1)]
\item classical data types (\texttt{bool}eans, \texttt{int}egers,
\texttt{long} integers, \texttt{float}ing point numbers, \texttt{str}ings)~;
\item standard containers (\texttt{list}s, \texttt{set}s, \texttt{tuple}s,
\texttt{dict}ionaries) ;
\item a few features or function inherited from
functional programming such as \texttt{list}/\texttt{set}/\texttt{dict}
comprehension and \texttt{reduce}, \texttt{map}, \texttt{filter} operators and
\item the \texttt{numpy} module that
provides a multi-dimensional array structure targeted at scientific computing. 
% "at scientific computing" looks strange for me but it may be my bad english...
\end{inparaenum}

@@

% mature?
Pythran is a mature \ac{OSS} available through its source code
repository (see \url{https://githib.com/serge-sans-paille/pythran}), the
Python Package Index (see \url{https://pypi.python.org/pypi/pythran}) and
a Debian repository (see \url{http://ridee.enstb.org/debian}). It can be
used as a source-to-source compiler infrastructure for a subset of the Python
language, a C++11 meta-programs generator or a native module generators. Each
of this target corresponds to a layer in the tool architecture described in
Figure~\ref{fig:archi}.

@@

% It looks like pythran generate a *.py file as an output. What did you mean with
% this arrow?
\begin{figure}[ht]
    \centering
    \begin{tikzpicture}[
            file/.style={align=center, node distance=0.4cm},
            state/.style={draw=black!50,fill=black!5,rectangle, align=center, node distance=0.4cm},
            box/.style={draw=black!50,fill=black!5,rectangle, rounded corners, drop shadow, align=center,
            node distance=0.4cm, inner sep=.2cm},
            mbox/.style={draw=black!50,fill=black!5,rectangle, rounded corners, drop shadow, align=center,
            node distance=0.4cm, inner sep=.2cm},
        ]
        \node[file] (python)                      {\textbf{Python Module [\texttt{.py}]}};
        \node[state] (ast)       [below=of python]       {\ac{AST}};
        \node[state] (ir)        [below=of ast]          {\ac{IR}};
        \node[state] (ir2)       [below=of ir]           {Optimized \ac{IR}};
        \node[state] (be0)       [xshift=1.5cm, below=of ir2]          {Python};
        \node[state] (be1)       [xshift=-1.5cm,below=of ir2]          {C++};
        \begin{pgfonlayer}{background} 
            \node[box]  (pythran)   [fit=(ast) (ir) (be0) (be1)]  {};
        \end{pgfonlayer}
        \node[mbox] (cxx)        [yshift=-.1cm, below=of be1]          {\texttt{cxx}};
        \node[file] (so)        [below=of cxx]          {\textbf{Native[\texttt{.so}]}};
        \node[file] (py)        [yshift=-.1cm, below=of be0]          {\textbf{Python[\texttt{.py}]}};

>>>>

Pythran\cite{pythran2013} is an \ac{AOT} compiler that focuses on the
optimization of the functional and procedural paradigms offered by Python,
leaving aside dynamicity and most \ac{OO} features to focus on scientific python modules.  Indeed, the Python
language and its standard library offers comprehensive services to the
programmers, but the language proves to be useful even with a limited subset of
features:
\begin{inparaenum}[1)]
\item classical data types (\texttt{bool}eans, \texttt{int}egers,
\texttt{long} integers, \texttt{float}ing point numbers, \texttt{str}ings)~;
\item standard containers (\texttt{list}s, \texttt{set}s, \texttt{tuple}s,
\texttt{dict}ionaries) ;
\item a few features or function inherited from
functional programming such as \texttt{list}/\texttt{set}/\texttt{dict}
comprehension and \texttt{reduce}, \texttt{map}, \texttt{filter} operators and
\item the \texttt{numpy} module that
provides a multi-dimensional array structure targeted at scientific computing. 
% "at scientific computing" looks strange for me but it may be my bad english...
\end{inparaenum}

@@

% mature?
% yes, mature! 1500+ validation test looks mature enough to me ;-)
Pythran is an \ac{OSS} available through its source code
repository (see \url{https://githib.com/serge-sans-paille/pythran}), the
Python Package Index (see \url{https://pypi.python.org/pypi/pythran}) and
a Debian repository (see \url{http://ridee.enstb.org/debian}). It can be
used as a source-to-source compiler infrastructure for a subset of the Python
language, a C++11 meta-programs generator or a native module generators. Each
of this target corresponds to a layer in the tool architecture described in
Figure~\ref{fig:archi}.

@@

% It looks like pythran generate a *.py file as an output. What did you mean with
% this arrow?
% SG: Pythran can generate a.py file if using the Python backend
\begin{figure}[ht]
    \centering
    \begin{tikzpicture}[
            file/.style={align=center, node distance=0.4cm},
            state/.style={draw=black!50,fill=black!5,rectangle, align=center, node distance=0.4cm},
            box/.style={draw=black!50,fill=black!5,rectangle, rounded corners, drop shadow, align=center,
            node distance=0.4cm, inner sep=.2cm},
            mbox/.style={draw=black!50,fill=black!5,rectangle, rounded corners, drop shadow, align=center,
            node distance=0.4cm, inner sep=.2cm},
        ]
        \node[file] (python)                      {\textbf{Python Module [\texttt{.py}]}};
        \node[state] (ast)       [below=of python]       {\ac{AST}};
        \node[state] (ir)        [below=of ast]          {\ac{IR}};
        \node[state] (ir2)       [below=of ir]           {Optimized \ac{IR}};
        \node[state] (be0)       [xshift=1.5cm, below=of ir2]          {Python};
        \node[state] (be1)       [xshift=-1.5cm,below=of ir2]          {C++};
        \begin{pgfonlayer}{background} 
            \node[box]  (pythran)   [fit=(ast) (ir) (be0) (be1)]  {};
        \end{pgfonlayer}
        \node[mbox] (cxx)        [yshift=-.1cm, below=of be1]          {\texttt{cxx}};
        \node[file] (so)        [below=of cxx]          {\textbf{Native[\texttt{.so}]}};
        \node[file] (py)        [yshift=-.1cm, below=of be0]          {\textbf{Python[\texttt{.py}]}};
